
\begin{thebibliography}{50}

\bibitem{PLAZA-site}
Website: https://bioinformatics.psb.ugent.be/plaza/, accessed on 24/04/2020

\bibitem{PLAZA-paper-1}
Van Bel, M., Diels, T., Vancaester, E., Kreft, L., Botzki, A., Van de Peer, Y., Coppens, F., Vandepoele, K. (2017) PLAZA 4.0: an integrative resource for functional, evolutionary and comparative plant genomics Nucleic Acids Res.

\bibitem{PLAZA-paper-2}
Vandepoele, K., Van Bel, M., Richard, G., Van Landeghem, S., Verhelst, B., Moreau, H., Van de Peer, Y., Grimsley, N., Piganeau, G. (2013) pico-PLAZA, a genome database of microbial photosynthetic eukaryotes. Environmental Microbiology 15(8):2147-53

\bibitem{ADVI}
Kucukelbir, Alp \& Tran, Dustin \& Ranganath, Rajesh \& Gelman, Andrew \& Blei, David. (2016). Automatic Differentiation Variational Inference. 18. 

\bibitem{ADVI-stan}
Kucukelbir, A., Tran, D., Ranganath, R., Gelman, A., \& Blei, D. M. (2017). Automatic differentiation variational inference. The Journal of Machine Learning Research, 18(1), 430-474.

\bibitem{vi-review}
Blei, D. M., Kucukelbir, A., \& Mcauliffe, J. D. (2017). Variational Inference: A Review for Statisticians. Journal of the American Statistical Association, 112(518), 859-877. doi:10.1080/01621459.2017.1285773

\bibitem{svi-cat}
Dang, T., \& Kishino, H. (2018). Stochastic Variational Inference for Bayesian Phylogenetics: A Case of CAT Model. doi:10.1101/358747

\bibitem{BBVI}
Ranganath, R., Gerrish, S., \& Blei, D. (2014, April). Black box variational inference. In Artificial Intelligence and Statistics (pp. 814-822).

\bibitem{SVGD-website}
Website: http://www.cs.utexas.edu/~qlearning/project.html?p=svgd, accessed on 26/07/2020

\bibitem{SVGD}
Liu, Q., \& Wang, D. (2016). Stein variational gradient descent: A general purpose bayesian inference algorithm. In Advances in neural information processing systems (pp. 2378-2386).

\bibitem{KL}
Kullback, S.; Leibler, R. A. On Information and Sufficiency. Ann. Math. Statist. 22 (1951), no. 1, 79--86. doi:10.1214/aoms/1177729694. https://projecteuclid.org/euclid.aoms/1177729694

\bibitem{ADAM}
Kingma, D. P., \& Ba, J. (2014). Adam: A method for stochastic optimization. arXiv preprint arXiv:1412.6980.

\bibitem{Flux}
Innes, M., Saba, E., Fischer, K., Gandhi, D., Rudilosso, M. C., Joy, N. M., ... & Shah, V. (2018). Fashionable modelling with Flux. arXiv preprint arXiv:1811.01457.

\bibitem{Stan}
Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B., Betancourt, M., ... & Riddell, A. (2017). Stan: A probabilistic programming language. Journal of statistical software, 76(1).

\bibitem{rice}
Stein, J.C., Yu, Y., Copetti, D. et al. Genomes of 13 domesticated and wild rice relatives highlight genetic conservation, turnover and innovation across the genus Oryza. Nat Genet 50, 285–296 (2018). https://doi.org/10.1038/s41588-018-0040-0

\bibitem{pymc3}
Salvatier J., Wiecki T.V., Fonnesbeck C. (2016) Probabilistic programming in Python using PyMC3. PeerJ Computer Science 2:e55 DOI: 10.7717/peerj-cs.55.

\bibitem{rice-split}
Thiel, T., Graner, A., Waugh, R., Grosse, I., Close, T. J., \& Stein, N. (2009). Evidence and evolutionary analysis of ancient whole-genome duplication in barley predating the divergence from rice. BMC evolutionary biology, 9, 209. https://doi.org/10.1186/1471-2148-9-209

\bibitem{AD}
Margossian, C. C. (2019). A review of automatic differentiation and its efficient implementation. Wiley Interdisciplinary Reviews: Data Mining and Knowledge Discovery, 9(4), e1305.

\bibitem{forwarddiff}
Revels, J., Lubin, M., & Papamarkou, T. (2016). Forward-mode automatic differentiation in Julia. arXiv preprint arXiv:1607.07892.

\bibitem{rao-blackwell}
Casella, G., & Robert, C. P. (1996). Rao-Blackwellisation of sampling schemes. Biometrika, 83(1), 81-94.

\bibitem{control-variate}
S. M. Ross. Simulation. Elsevier, 2002.

\bibitem{itol}
Ivica Letunic, Peer Bork, Interactive Tree Of Life (iTOL) v4: recent updates and new developments, Nucleic Acids Research, Volume 47, Issue W1, 02 July 2019, Pages W256–W259.

\bibitem{Turing}
Hong Ge, Kai Xu, Zoubin Ghahramani; Proceedings of the Twenty-First International Conference on Artificial Intelligence and Statistics, PMLR 84:1682-1690, 2018.

\bibitem{mrbayes}
Huelsenbeck, J. P., & Ronquist, F. (2001). MRBAYES: Bayesian inference of phylogenetic trees. Bioinformatics, 17(8), 754-755.

\bibitem{vi}
Beal, M. J. (2003). Variational algorithms for approximate Bayesian inference (Doctoral dissertation, UCL (University College London)).

\bibitem{bayes-ecology}
Ellison, A. M. (2004). Bayesian inference in ecology. Ecology letters, 7(6), 509-520.

\bibitem{vi-genome-environment}
Montesinos-López, O. A., Montesinos-López, A., Crossa, J., Montesinos-López, J. C., Luna-Vázquez, F. J., Salinas-Ruiz, J., Herrera-Morales, J. R., \& Buenrostro-Mariscal, R. (2017). A Variational Bayes Genomic-Enabled Prediction Model with Genotype × Environment Interaction. G3 (Bethesda, Md.), 7(6), 1833–1853. https://doi.org/10.1534/g3.117.041202

\bibitem{phylostan}
Fourment, M., & Darling, A. E. (2019). Evaluating probabilistic programming and fast variational Bayesian inference in phylogenetics. PeerJ, 7, e8272.

\bibitem{cat}
Dang, T., & Kishino, H. (2019). Stochastic variational inference for Bayesian phylogenetics: a case of CAT model. Molecular biology and evolution, 36(4), 825-833.

\bibitem{bbvi-advances}
Agrawal, A., Sheldon, D., & Domke, J. (2020). Advances in Black-Box VI: Normalizing Flows, Importance Weighting, and Optimization. arXiv preprint arXiv:2006.10343.

\bibitem{ncbi}
Sayers EW, Cavanaugh M, Clark K, Ostell J, Pruitt KD, Karsch-Mizrachi I. GenBank. Nucleic Acids Res. 2019 Jan 8;47(D1):D94-D99.

\bibitem{19}
Fourment, M., Magee, A. F., Whidden, C., Bilge, A., Matsen IV, F. A., \& Minin, V. N. (2020). 19 dubious ways to compute the marginal likelihood of a phylogenetic tree topology. Systematic Biology, 69(2), 209-220.


\end{thebibliography}